---
title: "A Comparison of Clustering Methods Using the Silhouette Score"
subtitle: "Jack Leary"
author: 
  - "University of North Carolina at Chapel Hill - Lineberger Comprehensive Cancer Center"
  - "University of Florida - Department of Biostatistics"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: yeti
    highlight: tango
    code_folding: show
    code_download: true
    toc: true
    toc_float:
      collpased: false
    df_print: kable
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE, 
                      fig.align = "center")
set.seed(629)
reticulate::use_virtualenv("~/Desktop/Python/science/venv/", required = TRUE)
```

# Libraries

## R

```{r}
library(dplyr)           # tidy data 
library(Seurat)          # single cell infrastructure
library(ggplot2)         # plots
library(cluster)         # clustering tools
library(magrittr)        # piping
library(SCISSORS)        # reclustering  
library(ggsignif)        # significance bars
library(paletteer)       # advanced colors
library(reticulate)      # Python interface
library(SeuratData)      # PBMC3k dataset
library(SAFEclustering)  # the SAFE algorithm
```

## Import libraries

```{python}
import anndata           # annotated single cell data
import numpy as np       # matrix algebra
import pandas as pd      # DataFrames
import scanpy as sc      # ScanPy
import giniclust3 as gc  # GiniClust3
```

# Data 

First we'll read in both our `SCISSORS`-processed PBMC3k dataset as well as the same data [as processed by the Satija Lab](https://satijalab.org/seurat/articles/pbmc3k_tutorial.html). We also load in a blank version of the object and pre-process it exactly as we did the `SCISSORS`-processed object. 

```{r}
pbmc_SCISSORS <- readRDS("~/Desktop/Data/pbmc3k.Rds")
sils_SCISSORS <- ComputeSilhouetteScores(pbmc_SCISSORS, avg = FALSE)
pbmc_Seurat <- LoadData("pbmc3k") %>% 
               subset(subset = nFeature_RNA < 2500) %>% 
               NormalizeData(verbose = FALSE) %>% 
               FindVariableFeatures(selection.method = "vst", nfeatures = 2000, verbose = FALSE) %>% 
               ScaleData(verbose = FALSE) %>% 
               RunPCA(verbose = FALSE) %>% 
               FindNeighbors(dims = 1:10, verbose = FALSE) %>% 
               FindClusters(resolution = 0.5, verbose = FALSE) %>% 
               RunTSNE(reduction = "pca", dims = 1:10, dim.embed = 2, verbose = FALSE)
pbmc_Seurat@reductions$fitsne <- pbmc_SCISSORS@reductions$tsne
pbmc <- LoadData("pbmc3k") %>% 
        subset(subset = nFeature_RNA < 2500) %>% 
        PrepareData(n.HVG = 4000, 
                    regress.cc = FALSE, 
                    regress.mt = FALSE, 
                    n.PC = 15, 
                    which.dim.reduc = "tsne",
                    initial.resolution = .4, 
                    random.seed = 629)
pbmc@reductions$fitsne <- pbmc_SCISSORS@reductions$tsne
sils_satija <- ComputeSilhouetteScores(pbmc_Seurat, avg = FALSE)
```

```{r}
pbmc_raw <- LoadData("pbmc3k") %>% subset(subset = nFeature_RNA < 2500) 
system.time(
  mk <- MultiK(seu = pbmc_raw, seed = 312)
)
pbmc_raw@reductions$fitsne <- pbmc_SCISSORS@reductions$tsne
```

```{r}
DiagMultiKPlot(mk$k, mk$consensus)
```

```{r}
clusters_k3 <- getClusters(pbmc_raw, optK = 3)
clusters_k7 <- getClusters(pbmc_raw, optK = 7)
clusters_k9 <- getClusters(pbmc_raw, optK = 9)
clusters_k10 <- getClusters(pbmc_raw, optK = 10)
pbmc_raw <- AddMetaData(pbmc_raw, metadata = clusters_k3$clusters, col.name = "MultiK_k3")
pbmc_raw <- AddMetaData(pbmc_raw, metadata = clusters_k7$clusters, col.name = "MultiK_k7")
pbmc_raw <- AddMetaData(pbmc_raw, metadata = clusters_k9$clusters, col.name = "MultiK_k9")
pbmc_raw <- AddMetaData(pbmc_raw, metadata = clusters_k10$clusters, col.name = "MultiK_k10")
pbmc_raw@reductions$fitsne <- pbmc_SCISSORS@reductions$tsne
```

```{r}
DimPlot(pbmc_raw, reduction = "fitsne", group.by = "MultiK_k3", cols = paletteer_d("ggsci::category20_d3"))
DimPlot(pbmc_raw, reduction = "fitsne", group.by = "MultiK_k7", cols = paletteer_d("ggsci::category20_d3"))
DimPlot(pbmc_raw, reduction = "fitsne", group.by = "MultiK_k9", cols = paletteer_d("ggsci::category20_d3"))
DimPlot(pbmc_raw, reduction = "fitsne", group.by = "MultiK_k10", cols = paletteer_d("ggsci::category20_d3"))
```

```{r}
pbmc_raw <- pbmc_raw %>% 
            NormalizeData(verbose = FALSE) %>% 
            FindVariableFeatures(selection.method = "vst", nfeatures = 2000, verbose = FALSE) %>% 
            ScaleData(verbose = FALSE) %>% 
            RunPCA(verbose = FALSE)
Idents(pbmc_raw) <- "MultiK_k10"
multik_k10_markers <- FindAllMarkers(pbmc_raw,
                                     test.use = "wilcox", 
                                     only.pos = TRUE, 
                                     verbose = FALSE, 
                                     random.seed = 629)
multik_k10_top5 <- multik_k10_markers %>% 
                   group_by(cluster) %>% 
                   slice_head(n = 5)
DotPlot(pbmc_raw, features = unique(multik_k10_top5$gene)) + 
  theme(axis.text.x = element_text(angle = 90, size = 9))
```

Here's what the default Satija Lab clustering looks like: 

```{r, fig.width=5, fig.height=3}
p0a <- DimPlot(pbmc_Seurat, reduction = "fitsne", group.by = "seurat_clusters", 
               pt.size = 1.5, cols = paletteer_d("ggsci::category20_d3")) + 
       labs(x = NULL, y = NULL, title = NULL) + 
       theme_yehlab() + 
       theme(panel.border = element_blank(), legend.text = element_text(size = 20)) + 
       guides(color = guide_legend(nrow = 1, override.aes = list(size = 8)))
p0a
```

And our results: 

```{r, fig.width=5, fig.height=3}
p0b <- DimPlot(pbmc_SCISSORS, pt.size = 1.5, group.by = "seurat_clusters", 
               reduction = "tsne", cols = paletteer_d("ggsci::category20_d3")) + 
       labs(x = NULL, y = NULL, title = NULL) + 
       theme_yehlab() + 
       theme(panel.border = element_blank(), legend.text = element_text(size = 20)) + 
       guides(color = guide_legend(nrow = 1, override.aes = list(size = 8)))
p0b
```

# Analysis 

## SAFEclustering

```{r, results='hide'}
SAFE_res <- individual_clustering(inputTags = data.frame(GetAssayData(pbmc, slot = "counts")), 
                                  mt_filter = FALSE, 
                                  nGene_filter = FALSE, 
                                  SC3 = TRUE, 
                                  gene_filter = FALSE, 
                                  CIDR = FALSE, 
                                  nPC.seurat = 15,
                                  Seurat = TRUE, 
                                  tSNE = TRUE, 
                                  low.genes = 200, 
                                  SEED = 629)
clust_ensembl <- SAFE(cluster_results = SAFE_res, 
                      program.dir = "~/Downloads/SAFE", 
                      HGPA = TRUE, 
                      MCLA = TRUE, 
                      CSPA = TRUE, 
                      SEED = 629)
pbmc@meta.data$SAFE_clust <- as.factor(as.integer(clust_ensembl$optimal_clustering) - 1)
cos_dist <- CosineDist(input = Embeddings(pbmc, "pca"))
sils_safe <- silhouette(dist = cos_dist, x = as.integer(pbmc@meta.data$SAFE_clust) - 1)
```

Here's what the results from `SAFE` look like on our t-SNE embedding:

```{r, fig.width=5, fig.height=3}
p0c <- DimPlot(pbmc, pt.size = 1.5, group.by = "SAFE_clust", 
               reduction = "fitsne", cols = paletteer_d("ggsci::category20_d3")) + 
       labs(x = NULL, y = NULL, title = NULL) + 
       theme_yehlab() + 
       theme(panel.border = element_blank(), legend.text = element_text(size = 20)) + 
       guides(color = guide_legend(nrow = 1, override.aes = list(size = 8)))
p0c
```

# GiniClust3

Here we'll use the `GiniClust3` Python library to identify rare cell clusters. 

```{r}
counts <- GetAssayData(pbmc, assay = "SCT", slot = "counts") %>% 
          as.data.frame()
```

## Cluster Cells

We bring the cells into Python. 

```{python}
counts = r.counts 
adata = anndata.AnnData(X=counts.T)
sc.pp.normalize_per_cell(adata, counts_per_cell_after=1e4)
```

Next we process them using `GiniClust` and `FanoClust`, and the consensus between the two. 

```{python}
# GiniIndexClust
gc.gini.calGini(adata)
adataGini = gc.gini.clusterGini(adata, neighbors=7)
# FanoClust
gc.fano.calFano(adata)
adataFano = gc.fano.clusterFano(adata)
# consensus cluster
consensusCluster = {}
consensusCluster['giniCluster'] = np.array(adata.obs['rare'].values.tolist())
consensusCluster['fanoCluster'] = np.array(adata.obs['fano'].values.tolist())
gc.consensus.generateMtilde(consensusCluster)
gc.consensus.clusterMtilde(consensusCluster)
```

We bring the results back into R. 

```{r}
gini_clusts <- as.integer(py$consensusCluster$finalCluster)
pbmc@meta.data$gini_clust <- gini_clusts
sils_gini <- cluster::silhouette(dist = cos_dist, x = gini_clusts)
```

Here's what the `GiniClust3` results look like on our t-SNE embedding:

```{r, fig.width=5, fig.height=3}
p0d <- DimPlot(pbmc, pt.size = 1.5, group.by = "gini_clust", 
               reduction = "fitsne", cols = paletteer_d("ggsci::category20_d3")) + 
       labs(x = NULL, y = NULL, title = NULL) + 
       theme_yehlab() + 
       theme(panel.border = element_blank(), legend.text = element_text(size = 20)) + 
       guides(color = guide_legend(nrow = 1, override.aes = list(size = 8)))
p0d
```

# Compare Results

We build a dataframe of the results that we'll pass into `ggplot2`. 

```{r}
clust_sils <- data.frame(Method = rep(c("SCISSORS", "Seurat", "SAFE", "GiniClust3"), each = ncol(pbmc)), 
                         Silhouette = c(sils_SCISSORS$Score, 
                                        sils_satija$Score, 
                                        sils_safe[, 3], 
                                        sils_gini[, 3]))
clust_sils %<>% mutate(Method = factor(Method,levels = c("GiniClust3", "Seurat", "SAFE", "SCISSORS"))) 
```

We see that our method gives the highest median silhouette score over the `Seurat` vignette's results, `GiniClust3`, and `SAFE`; the differences between `SCISSORS` and the other methods are statistically significant

```{r, fig.height=4, fig.width=5}
p1 <- ggplot(clust_sils, aes(x = Method, y = Silhouette, fill = Method)) + 
      geom_violin(draw_quantiles = 0.5, color = "black", scale = "width", size = 1.2) +
      geom_signif(test = wilcox.test, map_signif_level = TRUE, step_increase = .12, size = 1.2, textsize = 8, 
                  comparisons = list(c("SCISSORS", "GiniClust3"), 
                                     c("SCISSORS", "Seurat"), 
                                     c("SCISSORS", "SAFE"))) + 
      scale_fill_paletteer_d("ggsci::default_locuszoom") + 
      scale_y_continuous(breaks = c(-1, -0.5, 0, 0.5, 1), limits = c(-1, 1.6)) + 
      labs(y = "Silhouette Score", x = "Method", fill = NULL) + 
      theme(panel.grid = element_blank(), 
            legend.position = "none", 
            panel.background = element_blank(), 
            panel.border = element_rect(fill = NA, size = 1), 
            axis.title.x = element_text(size = 24), 
            axis.title.y = element_text(size = 24), 
            axis.text.y = element_text(size = 20), 
            axis.text.x = element_text(size = 20))
p1
```

# Save Figures

```{r}
ggsave(p0a, 
       filename = "Seurat_Clusters.pdf", 
       device = "pdf", 
       units = "in",
       path = "~/Desktop/R/SCISSORS/vignettes/figures_pub/PBMC/Method Comparison", 
       height = 8, 
       width = 8) 
ggsave(p0b, 
       filename = "SCISSORS_Clusters.pdf", 
       device = "pdf", 
       units = "in",
       path = "~/Desktop/R/SCISSORS/vignettes/figures_pub/PBMC/Method Comparison", 
       height = 8, 
       width = 8) 
ggsave(p0c, 
       filename = "SAFE_Clusters.pdf", 
       device = "pdf", 
       units = "in",
       path = "~/Desktop/R/SCISSORS/vignettes/figures_pub/PBMC/Method Comparison", 
       height = 8, 
       width = 8) 
ggsave(p0d, 
       filename = "GiniClust3_Clusters.pdf", 
       device = "pdf", 
       units = "in",
       path = "~/Desktop/R/SCISSORS/vignettes/figures_pub/PBMC/Method Comparison", 
       height = 9.5, 
       width = 9.5) 
ggsave(p1, 
       filename = "Sil_Scores.pdf", 
       device = "pdf", 
       units = "in",
       path = "~/Desktop/R/SCISSORS/vignettes/figures_pub/PBMC/Method Comparison", 
       height = 8, 
       width = 8) 
```
