---
title: "Reclustering Cancer-Associated Fibroblasts Using SCISSORS"
subtitle: "Jack Leary"
author: 
  - "University of North Carolina at Chapel Hill - Lineberger Comprehensive Cancer Center"
  - "University of Florida - Department of Biostatistics"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: paper
    highlight: tango
    df_print: kable
    toc: true
    toc_float: true
    code_folding: show
    code_download: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE, 
                      results = "hold", 
                      fig.align = "center")
reticulate::use_virtualenv("~/Desktop/Python/science/venv/", required = TRUE)
```

# Introduction

In 2017, the Tuveson Lab at Cold Spring Harbor Cancer Center published a paper written by Elyada *et al* that detailed the discovery of cancer-associated fibroblasts (CAFs) in mice. The subtypes were then validated in human samples affected with PDAC in a subsequent paper released in 2019. Here we will use SCISSORS to identify the CAF subtypes within the larger stroma population, fine-grained immune cell types within broadly-defined immune clusters, and ductal & PDAC subtypes within the ductal group. You can install SCISSORS from [our GitHub repository](https://github.com/jr-leary7/SCISSORS).

# Libraries

## R

```{r, results='hide'}
library(VAM)           # single cell GSEA
library(dplyr)         # tidy data 
library(Seurat)        # single cell infrastructure
library(ggplot2)       # pretty plots
library(SingleR)       # cell type assignment
library(janitor)       # clean data
library(magrittr)      # pipes
library(decoderr)      # de novo deconvolution 
library(SCISSORS)      # our package
library(mixtools)      # Gaussian mixture model estimation
library(ggsignif)      # significance bars
library(patchwork)     # align plots
library(latex2exp)     # LaTeX
library(paletteer)     # color palettes
library(CONICSmat)     # CNV estimation
library(reticulate)    # Python interface
library(kableExtra)    # pretty tables
library(wesanderson)   # more color palettes
```

## Python

```{python}
import numpy as np
from openTSNE import TSNEEmbedding
from openTSNE import initialization
from openTSNE.affinity import Multiscale
from openTSNE.affinity import PerplexityBasedNN
```

# Data

```{r}
raw_counts <- Read10X(data.dir = "~/Desktop/Data/Tuveson_v6/count/raw_feature_bc_matrix/")
pdac <- CreateSeuratObject(raw_counts, 
                           project = "Elyada", 
                           min.cells = 3, 
                           min.features = 500)
pdac@meta.data %<>%
  mutate(sample = case_when(substr(rownames(.), 17, 18) == "-1" ~ "SRR9274536", 
                            substr(rownames(.), 17, 18) == "-2" ~ "SRR9274537", 
                            substr(rownames(.), 17, 18) == "-3" ~ "SRR9274538", 
                            substr(rownames(.), 17, 18) == "-4" ~ "SRR9274539", 
                            substr(rownames(.), 17, 18) == "-5" ~ "SRR9274541", 
                            substr(rownames(.), 17, 18) == "-6" ~ "SRR9274542", 
                            substr(rownames(.), 17, 18) == "-7" ~ "SRR9274543",
                            substr(rownames(.), 17, 18) == "-8" ~ "SRR9274544"), 
         sex = case_when(sample == "SRR9274536" ~ "F", 
                         sample == "SRR9274537" ~ "M", 
                         sample == "SRR9274538" ~ "M", 
                         sample == "SRR9274539" ~ "M", 
                         sample == "SRR9274541" ~ "F", 
                         sample == "SRR9274542" ~ "F", 
                         sample == "SRR9274543" ~ "M", 
                         sample == "SRR9274544" ~ "M"), 
         condition = case_when(sample == "SRR9274536" ~ "PDAC", 
                               sample == "SRR9274537" ~ "PDAC", 
                               sample == "SRR9274538" ~ "PDAC", 
                               sample == "SRR9274539" ~ "PDAC", 
                               sample == "SRR9274541" ~ "AdjNorm", 
                               sample == "SRR9274542" ~ "PDAC", 
                               sample == "SRR9274543" ~ "AdjNorm", 
                               sample == "SRR9274544" ~ "PDAC"))
```

```{r, echo=FALSE}
rm(raw_counts)
```

Next, we read in a dataset of basal-like and classical PDAC marker genes that we'll use later on to perform enrichment analysis. 

```{r}
load("~/Desktop/Data/cmbSubtypes.RData")
pdac_50_genes <- subtypeGeneList[[4]]
basal_genes <- pdac_50_genes[pdac_50_genes$BasalLike, ]$geneSymbol
classical_genes <- pdac_50_genes[!pdac_50_genes$BasalLike, ]$geneSymbol
```

# Preprocessing

We run the typical single cell pre-processing steps on our cells - normalization, dimension reduction, and clustering. 

```{r, warning=FALSE}
pdac <- PrepareData(seurat.object = pdac, 
                    n.HVG = 4000, 
                    n.PC = 20, 
                    regress.mt = FALSE, 
                    regress.cc = FALSE, 
                    which.dim.reduc = "tsne", 
                    initial.resolution = 0.4,
                    k.val = 125,  
                    random.seed = 629)
```

Let's check the default t-SNE embedding & Louvain clustering. 

```{r}
p0 <- DimPlot(pdac, reduction = "tsne", pt.size = 0.75) + 
      scale_color_paletteer_d("ggthemes::Classic_20") + 
      labs(x = "t-SNE 1", y = "t-SNE 2") + 
      theme_yehlab() + 
      theme(legend.text = element_text(size = 10)) + 
      guides(color = guide_legend(nrow = 1, override.aes = list(size = 4)))
p0
```

We compute the distribution of the silhouette scores for each cluster. 

```{r}
sil_df <- ComputeSilhouetteScores(pdac, avg = FALSE)
p1 <- ggplot(sil_df, aes(x = Cluster, y = Score, fill = Cluster)) + 
      geom_violin(draw_quantiles = .5, color = "black", scale = "width") + 
      scale_fill_paletteer_d("ggthemes::Classic_20") + 
      labs(y = "Silhouette Score", x = "Louvain Clusters", fill = NULL) + 
      theme_minimal() + 
      theme(panel.grid = element_blank(), 
            panel.border = element_rect(fill = NA, size = 1), 
            legend.position = "none", 
            axis.title.x = element_text(size = 22), 
            axis.title.y = element_text(size = 22), 
            axis.text = element_text(size = 16), 
            axis.ticks = element_line(), 
            plot.subtitle = element_text(face = "italic", size = 10))
```

Some of our clusters have less-than-ideal fits. 

```{r}
p1
```

Lastly, we'll identify marker genes for each of the clusters. 

```{r, fig.width=8, fig.height=4}
pdac_markers <- FindSpecificMarkers(pdac, de.method = "wilcox", log2fc.cutoff = 0.1)
top5_pdac_markers <- pdac_markers %>% 
                     group_by(cluster) %>% 
                     arrange(desc(avg_log2FC)) %>% 
                     slice_head(n = 5)
p2 <- DotPlot(pdac, features = unique(top5_pdac_markers$gene), dot.scale = 15) + 
      scale_color_gradientn(colors = paletteer_d("wesanderson::Zissou1")) +
      labs(color = "Expression", size = "% Expressed", y = "Louvain Cluster") + 
      theme(axis.text.x = element_text(angle = 90, size = 16, vjust = 0.5), 
            legend.position = "right", legend.justification = "center", 
            panel.border = element_rect(fill = NA, size = 1, color = "black"), 
            axis.line = element_blank(), 
            legend.title = element_text(size = 18), 
            axis.title.x = element_blank(), 
            axis.title.y = element_text(size = 20), 
            axis.text.y = element_text(size = 18)) + 
      guides(color = guide_colorbar(title.position = "top", barheight = unit(3, units = "cm"), title.hjust = 0.5), 
             size = guide_legend(title.position = "top", title.hjust = 0.5))
p2
```

## Fit-SNE

For information on how to install the `openTSNE` implementation of Fit-SNE and how to run the algorithm, please visit [the excellent GitHub repository of Pavlin Policar](https://github.com/pavlin-policar/openTSNE).

First we'll run a simple, standard Fit-SNE embedding. It's necessary to make the PCA embeddings accessible by Python.

```{r}
pc_df <- Embeddings(pdac, reduction = "pca")
```

```{python}
# import data
pc_df = np.array(r.pc_df)
# run Fit-SNE
affin = PerplexityBasedNN(pc_df, perplexity=30, metric='cosine', random_state=629)
init = initialization.pca(pc_df, random_state=629)
tsne1 = TSNEEmbedding(init, affin, negative_gradient_method='fft')
embed1 = tsne1.optimize(n_iter=350, exaggeration=12, momentum=0.6) 
embed2 = embed1.optimize(n_iter=750, momentum=0.8)
```

The embedding looks good, so we'll use it going forwards.

```{r}
embed <- as.matrix(py$embed2)
rownames(embed) <- colnames(pdac)
pdac@reductions$fitsne <- CreateDimReducObject(embeddings = embed, 
                                               key = "FitSNE_", 
                                               assay = "SCT", 
                                               global = TRUE)
p3 <- DimPlot(pdac, reduction = "fitsne", pt.size = 0.75) + 
      scale_color_paletteer_d("ggthemes::Classic_20") + 
      labs(x = "Fit-SNE 1", y = "Fit-SNE 2") + 
      theme_yehlab() +
      theme(legend.text = element_text(size = 10)) + 
      guides(color = guide_legend(nrow = 1, override.aes = list(size = 4)))
p3
```

Due to the way `Seurat` accesses cell embeddings, we'll need to replace our original t-SNE dimension reduction in our `Seurat` object with the new Fit-SNE version. We'll keep the original Barnes-Hut t-SNE embedding under a separate name.

```{r}
pdac@reductions$bh_tsne <- pdac@reductions$tsne
pdac@reductions$tsne <- pdac@reductions$fitsne
```

```{r, echo=FALSE, results='hide'}
rm(embed, pc_df)
gc(verbose = FALSE)
```

# SingleR Cell Type Identification

## Single Cell RNA-seq Reference Data

Here we use the `SingleR` package to identify broad cell types. The reference dataset we load is an `sctransform`-normalized version of the raw counts available in `celldex::BaronPancreasData()`, which consists of normal pancreas cells that were sequenced and annotated by the researchers.

```{r}
sc_ref <- readRDS("/Volumes/labs/Home/Jen Jen Yeh Lab/Jack/scRNAseq/Seurat/single_cell_ref_normalized.Rds")
sc_preds <- SingleR(test = data.frame(pdac@assays$SCT@data), 
                    ref = sc_ref, 
                    labels = sc_ref$label, 
                    clusters = pdac$seurat_clusters, 
                    de.method = "wilcox")
pdac[["SingleR.labels.sc"]] <- sc_preds$labels[match(pdac[[]][["seurat_clusters"]], rownames(sc_preds))]
pdac$SingleR.labels.sc <- case_when(pdac$SingleR.labels.sc == "acinar" ~ "Acinar", 
                                    pdac$SingleR.labels.sc == "activated_stellate" ~ "Activated Stellate",
                                    pdac$SingleR.labels.sc == "quiescent_stellate" ~ "Quiescent Stellate",
                                    pdac$SingleR.labels.sc == "ductal" ~ "Ductal", 
                                    pdac$SingleR.labels.sc == "macrophage" ~ "Macrophage",
                                    pdac$SingleR.labels.sc == "t_cell" ~ "T")
```

We can see that there's a large immune population, as well as smaller ductal, fibroblast (denoted activated stellate in the reference dataset), stromal (denoted quiescent stellate in the reference dataset), and acinar groups. These broad cell types line up with what we expected to see given the authors' original cell cluster annotations. 

```{r}
p4 <- DimPlot(pdac, reduction = "tsne", group.by = "SingleR.labels.sc", pt.size = 0.75) + 
      scale_color_manual(values = paletteer_d("ggsci::default_nejm")) + 
      labs(x = "Fit-SNE 1", y = "Fit-SNE 2") + 
      theme(plot.title = element_blank()) + 
      theme_yehlab() + 
      guides(color = guide_legend(nrow = 2, override.aes = list(size = 4)))
p4
```

## Bulk Tissue RNA-seq Reference Data

This dataset is composed of labeled & log-normalized bulk RNA-seq samples from the Human Primary Cell Atlas.

```{r}
bulk_ref <- celldex::HumanPrimaryCellAtlasData()
bulk_preds <- SingleR(test = data.frame(pdac@assays$SCT@data), 
                      ref = bulk_ref, 
                      labels = bulk_ref$label.main, 
                      clusters = pdac$seurat_clusters, 
                      de.method = "wilcox")
pdac[["SingleR.labels.bulk"]] <- bulk_preds$labels[match(pdac[[]][["seurat_clusters"]], rownames(bulk_preds))]
pdac$SingleR.labels.bulk <- case_when(pdac$SingleR.labels.bulk == "B_cell" ~ "B", 
                                      pdac$SingleR.labels.bulk == "Epithelial_cells" ~ "Epithelial", 
                                      pdac$SingleR.labels.bulk == "Endothelial_cells" ~ "Endothelial", 
                                      pdac$SingleR.labels.bulk == "MSC" ~ "Mesenchymal Stem Cell", 
                                      pdac$SingleR.labels.bulk == "Pre-B_cell_CD34-" ~ "Pre-B CD34-", 
                                      pdac$SingleR.labels.bulk == "T_cells" ~ "T",
                                      TRUE ~ pdac$SingleR.labels.bulk)
```

The bulk reference gives us somewhat more granular labels for the immune cells, and confirms the identities of the epithelial and stromal clusters. 

```{r}
p5 <- DimPlot(pdac, reduction = "tsne", group.by = "SingleR.labels.bulk", pt.size = 0.75) + 
      scale_color_manual(values = paletteer_d("ggsci::default_nejm")) + 
      labs(x = "Fit-SNE 1", y = "Fit-SNE 2") + 
      theme(plot.title = element_blank()) + 
      theme_yehlab() + 
      guides(color = guide_legend(nrow = 2, override.aes = list(size = 4)))
p5
```

```{r, echo=FALSE, results='hide'}
rm(sc_preds, bulk_preds, bulk_ref, sc_ref)
gc(verbose = FALSE)
```

One shouldn't use `SingleR` as the final authority for cell types, but we were able to confirm the identities of the ductal and fibroblast clusters, which is important for this dataset as the cells we are most interested in are the cancer-associated fibroblasts (CAFs). 

# CONICSmat CNV Estimation

Next we'll attempt to identify malignant cells using single-cell copy number variation estimation as implemented in the `CONCISmat` package. Details of the GMM methodology used can be found at [the Diaz Lab's GitHub repository](https://github.com/diazlab/CONICS). Note: this step is memory-intensive because 1) it requires the sparse counts matrix to be cast to an in-memory dense matrix and 2) a *lot* of Gaussian mixture models get estimated. If your machine doesn't have a lot of RAM it might be best to skip this and manually annotate the malignant cells instead through the usage of canonical markers, GSEA, etc. 

```{r, results="hide"}
chrom_regions <- read.table("/Volumes/labs/Home/Jen Jen Yeh Lab/Jack/scRNAseq/chrom_arm_positions.txt", 
                            sep = "\t", 
                            row.names = 1, 
                            header = TRUE)
gene_pos <- getGenePositions(rownames(pdac))
cpm <- t(t(as.matrix(pdac@assays$SCT@counts)) / colSums(as.matrix(pdac@assays$SCT@counts))) * 1e5
cpm <- log2(cpm + 1)
norm_factor <- calcNormFactors(cpm)
cnv_est <- plotAll(mat = cpm, 
                   normFactor = norm_factor, 
                   regions = chrom_regions, 
                   gene_pos = gene_pos, 
                   fname = "Elyada")
```

```{r, echo=FALSE, results='hide'}
gc(verbose = FALSE)
```

## Visualizing CNVs

After estimating CNVs, we cluster the cells into $k = 3$ clusters, with the hope of finding one large cluster of normal cells and two smaller clusters composed of CAFs and PDAC cells. 

```{r}
bic_table <- read.table("./Elyada_BIC_LR.txt", 
                        sep = "\t", 
                        row.names = 1, 
                        header = TRUE, 
                        check.names = FALSE)
cand_regions <- rownames(bic_table[bic_table$`BIC difference` > 1000 & bic_table$`LRT adj. p-val` < .01, ])
pdf("~/Desktop/R/SCISSORS/vignettes/figures_supp/Elyada/CONICSmat_Heatmap_new.pdf", width = 12, height = 6)
hist1 <- plotHistogram(cnv_est[, cand_regions], 
                       cpm, 
                       clusters = 3, 
                       zscoreThreshold = 3, 
                       celltypes = pdac$SingleR.labels.bulk, 
                       patients = pdac$sample)
dev.off()
```


```{r}
normal <- which(hist1 == 1)
malignant <- which(hist1 != 1)
pdac@meta.data$malig <- ifelse(rownames(pdac@meta.data) %in% names(normal), "Normal", "Malignant")
```

We add the normal vs. malignant cell labels in to our `Seurat` object's metadata, then visualize the results. As expected, the malignant cells are located in the clusters identified by `SingleR` as ductal cells and fibroblasts. This indicates that `CONICSmat` did a solid job of estimating the CNVs - no easy feat with sparse, noisy single-cell data. 

```{r}
p6 <- DimPlot(pdac, reduction = "tsne", group.by = "malig", pt.size = 0.75) + 
      scale_color_manual(values = wes_palette("Zissou1", n = 5)[c(5, 2)]) + 
      labs(x = "Fit-SNE 1", y = "Fit-SNE 2") + 
      theme(plot.title = element_blank()) + 
      theme_yehlab() + 
      guides(color = guide_legend(nrow = 1, override.aes = list(size = 4)))
p6
```

When looking at the `SingleR` labels from the bulk reference, we see that the Epithelial (Ductal) and Mesenchymal Stem Cell (Fibroblast) clusters have the highest proportions of malignant cells, which is exactly what we expected to see. 

```{r}
pdac@meta.data %>% 
  group_by(SingleR.labels.bulk) %>% 
  summarise(MeanMalig = mean(case_when(malig == "Malignant" ~ 1, TRUE ~ 0))) %>% 
  mutate(MeanMalig = formattable::percent(MeanMalig)) %>% 
  kbl(booktabs = TRUE, col.names = c("Bulk Reference SingleR Label", "% Malignant Cells")) %>% 
  kable_minimal(full_width = FALSE)
```

```{r, echo=FALSE, results="hide"}
rm(cpm, bic_table, chrom_regions, cnv_est, gene_pos, cand_regions, hist1, normal, norm_factor, malignant)
gc(verbose = FALSE)
```

# DECODER

Next, we use [Dr. Xianlu Peng's DECODER](https://github.com/laurapeng/decoderr) in order to deconvolve the dataset and assign weights to each cell using non-negative matrix factorization. We calculate basal & classical PDAC, normal & activated stroma, immune, and endocrine & exocrine pancreas compartment weights.

```{r}
sample_wts_unscaled <- Decon_single_sample(refSet = "TCGA_RNAseq_PAAD", 
                                           data = pdac@assays$SCT@data, 
                                           geneIDType = "geneSymbol")
sample_wts <- Norm_PDAC_weights(sample_wts_unscaled)
pdac <- AddMetaData(pdac, sample_wts$Immune, "immune")
pdac <- AddMetaData(pdac, sample_wts$bcRatio, "bc_ratio")
pdac <- AddMetaData(pdac, sample_wts$Exocrine, "exocrine")
pdac <- AddMetaData(pdac, sample_wts$Endocrine, "endocrine")
pdac <- AddMetaData(pdac, sample_wts$BasalTumor, "basal_like")
pdac <- AddMetaData(pdac, sample_wts$ClassicalTumor, "classical")
pdac <- AddMetaData(pdac, sample_wts$NormalStroma, "norm_stroma")
pdac <- AddMetaData(pdac, sample_wts$ActivatedStroma, "act_stroma")
```

## Visualizing DECODER Weights

### Basal PDAC

The basal weights are highest in a subcluster of the ductal group identified through `SingleR`. This is interesting as the authors did not find evidence of the basal subtype in their paper. 

```{r}
p7 <- FeaturePlot(pdac, reduction = "tsne", features = "basal_like", pt.size = 0.75) + 
      scale_color_gradientn(colors = wesanderson::wes_palette("Zissou1")) + 
      labs(x = "Fit-SNE 1", y = "Fit-SNE 2") + 
      theme(plot.title = element_blank()) + 
      theme_yehlab() + 
      NoLegend()
p7
```

### Classical PDAC

The classical weights are highest in another subcluster of the ductal cluster, and cells with high classical weights do not collocate with those that have high basal weights. The putative classical and basal PDAC cells also align closely with the cells identified through `CONCISmat` as malignant. 

```{r}
p8 <- FeaturePlot(pdac, reduction = "tsne", features = "classical", pt.size = 0.75) + 
      scale_color_gradientn(colors = wesanderson::wes_palette("Zissou1")) + 
      labs(x = "Fit-SNE 1", y = "Fit-SNE 2") + 
      theme(plot.title = element_blank()) + 
      theme_yehlab() + 
      NoLegend()
p8
```

### Exocrine Pancreas

The cluster identified through `SingleR` as acinar cells is the only cluster with high exocrine pancreas weights. 

```{r}
p9 <- FeaturePlot(pdac, reduction = "tsne", features = "exocrine", pt.size = 0.75) +
      scale_color_gradientn(colors = wesanderson::wes_palette("Zissou1")) + 
      labs(x = "Fit-SNE 1", y = "Fit-SNE 2") + 
      theme(plot.title = element_blank()) + 
      theme_yehlab() + 
      NoLegend()
p9
```

### Endocrine Pancreas

No cells have high endocrine pancreas weights. 

```{r}
p10 <- FeaturePlot(pdac, reduction = "tsne", features = "endocrine", pt.size = 0.75) +
       scale_color_gradientn(colors = wesanderson::wes_palette("Zissou1")) + 
       labs(x = "Fit-SNE 1", y = "Fit-SNE 2") + 
       theme(plot.title = element_blank()) + 
       theme_yehlab() + 
       NoLegend()
p10
```

### Immune

Once again, we confirm the largeness of the immune cell population in this dataset. 

```{r}
p11 <- FeaturePlot(pdac, reduction = "tsne", features = "immune", pt.size = 0.75) + 
       scale_color_gradientn(colors = wesanderson::wes_palette("Zissou1")) + 
       labs(x = "Fit-SNE 1", y = "Fit-SNE 2") + 
       theme(plot.title = element_blank()) + 
       theme_yehlab() + 
       NoLegend()
p11
```

### Normal Stroma

Cells with high normal stroma stroma weights are located in the cluster identified by `SingleR` as being stromal cells. 

```{r}
p12 <- FeaturePlot(pdac, reduction = "tsne", features = "norm_stroma", pt.size = 0.75) +
       scale_color_gradientn(colors = wesanderson::wes_palette("Zissou1")) + 
       labs(x = "Fit-SNE 1", y = "Fit-SNE 2") + 
       theme(plot.title = element_blank()) + 
       theme_yehlab() + 
       NoLegend()
p12
```

### Activated Stroma

Cells with high activated stroma weights are also located in the fibroblast cluster, and do not intersect with the cells that have high normal stroma scores. This indicates that `SCISSORS` will most likely perform well on the fibroblast cluster and be able to quickly tease out the cell subtypes.

```{r}
p13 <- FeaturePlot(pdac, reduction = "tsne", features = "act_stroma", pt.size = 0.75) +
       scale_color_gradientn(colors = wesanderson::wes_palette("Zissou1")) + 
       labs(x = "Fit-SNE 1", y = "Fit-SNE 2") + 
       theme(plot.title = element_blank()) + 
       theme_yehlab() + 
       NoLegend()
p13
```

# SCISSORS

## Fibroblasts

The fibroblast marker genes provided by Elyada *et al* match the `SingleR` results defining cluster 6 as containing fibroblasts. 

```{r}
p14 <- FeaturePlot(pdac, reduction = "tsne", features = "COL1A1", pt.size = 0.75) +
       scale_color_gradientn(colors = wesanderson::wes_palette("Zissou1")) + 
       labs(title = "COL1A1") + 
       theme_yehlab() + 
       NoLegend() + 
       theme(axis.title = element_blank())
p15 <- FeaturePlot(pdac, reduction = "tsne", features = "COL3A1", pt.size = 0.75) +
       scale_color_gradientn(colors = wesanderson::wes_palette("Zissou1")) + 
       labs(title = "COL3A1") + 
       theme_yehlab() + 
       NoLegend() + 
       theme(axis.title = element_blank())
p16 <- FeaturePlot(pdac, reduction = "tsne", features = "LUM", pt.size = 0.75) + 
       scale_color_gradientn(colors = wesanderson::wes_palette("Zissou1")) + 
       labs(title = "LUM") + 
       theme_yehlab() + 
       NoLegend() + 
       theme(axis.title = element_blank())
p17 <- FeaturePlot(pdac, reduction = "tsne", features = "DCN", pt.size = 0.75) + 
       scale_color_gradientn(colors = wesanderson::wes_palette("Zissou1")) + 
       labs(title = "DCN") + 
       theme_yehlab() + 
       NoLegend() + 
       theme(axis.title = element_blank())
(p14 | p15) / (p16 | p17)
```

Here's the cells we'll be reclustering. 

```{r}
fibro_cells <- rownames(pdac@meta.data[pdac@meta.data$seurat_clusters == 8, ])
p18 <- DimPlot(pdac, reduction = "tsne", cells.highlight = fibro_cells, cols.highlight = "navy", pt.size = 0.75) + 
       labs(x = NULL, y = NULL) + 
       theme_yehlab() + 
       NoLegend()
p18 / p3
```

### Recluster Fibroblasts

Here we use `ReclusterCells()` to identify subclusters within cluster 6. We find five distinct subclusters.

```{r, warning=FALSE}
fibro <- ReclusterCells(pdac,
                        which.clust = c(8, 10), 
                        merge.clusters = TRUE,  
                        n.HVG = 4000, 
                        n.PC = 20, 
                        resolution.vals = c(.05, .1, .2, .3, .4), 
                        k.vals = c(10, 20, 30, 40), 
                        redo.embedding = TRUE)
fibro_pc <- Embeddings(fibro, "pca")
```

We'll again run Fit-SNE on the reclustered cells, for consistencies sake. 

```{python}
# import data
fibro_pc = r.fibro_pc
# run Fit-SNE
affin_fibro = PerplexityBasedNN(fibro_pc, perplexity=50, metric='cosine', random_state=629)
init = initialization.pca(fibro_pc, random_state=629)
tsne_f = TSNEEmbedding(init, affin_fibro, negative_gradient_method='fft')
embed_f1 = tsne_f.optimize(n_iter=150, exaggeration=2, momentum=0.6) 
embed_f2 = embed_f1.optimize(n_iter=750, exaggeration=1, momentum=0.8)
```

Pulling the results back into R and visualizing them shows clear separation between the subclusters. There's some noise in subcluster 0, but other than that the re-embedding looks solid. 

```{r}
embed_fibro <- as.matrix(py$embed_f2)
rownames(embed_fibro) <- colnames(fibro)
fibro@reductions$bh_tsne <- fibro@reductions$tsne
fibro@reductions$tsne<- CreateDimReducObject(embeddings = embed_fibro, 
                                             key = "FitSNE_", 
                                             assay = "SCT", 
                                             global = TRUE)
p19 <- DimPlot(fibro, reduction = "tsne", pt.size = 1.5) + 
       scale_color_manual(values = paletteer_d("ggsci::default_nejm")) + 
       labs(x = "Fit-SNE 1", y = "Fit-SNE 2") + 
       theme_yehlab() + 
       guides(color = guide_legend(nrow = 1, override.aes = list(size = 4)))
p19
```


```{r}
fibro_markers <- FindAllMarkers(fibro, 
                                logfc.threshold = 0.25, 
                                test.use = "wilcox", 
                                only.pos = TRUE, 
                                verbose = FALSE, 
                                random.seed = 312) %>% 
                 filter(p_val_adj < 0.05)
DotPlot(fibro, 
        features = (fibro_markers %>% group_by(cluster) %>% arrange(desc(avg_log2FC)) %>% slice_head(n = 5) %>% pull(gene))) + 
  theme(axis.text.x = element_text(angle = 90))
```

### Cell Type Identification

#### Normal Stroma Cells

First we identify the endothelial and perivascular cells in clusters 3 & 1 using PLVAP and RGS5 expression. 

```{r}
p20 <- FeaturePlot(fibro, reduction = "tsne", features = "PLVAP", pt.size = 1.5) +
       scale_color_gradientn(colors = wesanderson::wes_palette("Zissou1")) + 
       labs(title = "PLVAP") + 
       theme_yehlab() + 
       NoLegend() + 
       theme(axis.title = element_blank())
p21 <- FeaturePlot(fibro, reduction = "tsne", features = "RGS5", pt.size = 1.5) + 
       scale_color_gradientn(colors = wesanderson::wes_palette("Zissou1")) + 
       labs(title = "RGS5") + 
       theme_yehlab() + 
       NoLegend() + 
       theme(axis.title = element_blank())
(p20 | p21) / p19
```

#### CAFs

Next we'll use `VAM` to perform GSEA on the iCAF, myCAF, and apCAF gene sets. 

```{r}
icaf_genes <- c("IL6", "PDGFRA", "CXCL12", "CFD", "LMNA", "AGTR1", 
                "HAS1", "CXCL1", "CXCL2", "CCL2", "CXCL8")  # use CXCL8 instead of IL8
mycaf_genes <- c("ACTA2", "TAGLN", "MMP11", "MYL9", "HOPX", "POSTN", "TPM1", "TPM2")
apcaf_genes_mouse <- c("Slpi", "Ptgis", "Nkain4", "Hspb1", "Cav1", "Arhgdib", "Lgals7", "Spint2", 
                       "Dmkn", "Saa3", "Clu", "Slc9a3r1", "Krt8", "Cxadr", "Ezr", "H2-Ab1", "Cd74", "Fth1")
apcaf_genes_human <- ConvertGeneOrthologs(gene.vec = apcaf_genes_mouse)
pan_caf_genes <- c("COL1A1", "FAP", "PDPN", "DCN", "VIM")
```

```{r}
gene_sets <- list(icaf_genes, mycaf_genes, apcaf_genes, pan_caf_genes)
names(gene_sets) <- c("iCAF", "myCAF", "apCAF", "Pan-CAF")
for (i in seq(gene_sets)) {
  gene_sets[[i]] <- gene_sets[[i]][gene_sets[[i]] %in% rownames(fibro)]
}
fibro <- vamForSeurat(fibro, 
                      gene.set.collection = gene_sets, 
                      gamma = TRUE)
DefaultAssay(fibro) <- "VAMcdf"
```

Let's plot the results. Clusters 0, 2, and 4 are composed of iCAF, myCAF, and apCAF cells, respectively. 

```{r}
p22 <- FeaturePlot(fibro, reduction = "tsne", features = "iCAF", pt.size = 1.5) +
       scale_color_gradientn(colors = wesanderson::wes_palette("Zissou1")) + 
       labs(title = "iCAF") + 
       theme_yehlab() + 
       NoLegend() + 
       theme(axis.title = element_blank())
p23 <- FeaturePlot(fibro, reduction = "tsne", features = "myCAF", pt.size = 1.5) + 
       scale_color_gradientn(colors = wesanderson::wes_palette("Zissou1")) + 
       labs(title = "myCAF") + 
       theme_yehlab() + 
       NoLegend() + 
       theme(axis.title = element_blank())
p24 <- FeaturePlot(fibro, reduction = "tsne", features = "apCAF", pt.size = 1.5) + 
       scale_color_gradientn(colors = wesanderson::wes_palette("Zissou1")) + 
       labs(title = "apCAF") + 
       theme_yehlab() + 
       NoLegend() + 
       theme(axis.title = element_blank())
(p22 | p23 | p24) / p19
```

Here's the final cell labels for all of our stromal cells:

```{r}
fibro$label <- case_when(fibro$seurat_clusters == 0 ~ "myCAF", 
                         fibro$seurat_clusters == 1 ~ "Perivascular", 
                         fibro$seurat_clusters == 2 ~ "iCAF", 
                         fibro$seurat_clusters == 3 ~ "Endothelial", 
                         fibro$seurat_clusters == 4 ~ "apCAF")
Idents(fibro) <- "label"
p25 <- DimPlot(fibro, reduction = "tsne", pt.size = 1.5) + 
       scale_color_manual(values = paletteer_d("ggsci::default_nejm")) + 
       labs(x = "Fit-SNE 1", y = "Fit-SNE 2") + 
       theme_yehlab() + 
       theme(plot.title = element_blank()) + 
       guides(color = guide_legend(nrow = 1, override.aes = list(size = 4)))
p25
```

### CAF-only Visualization

Here we'll generate violin plots for each of the CAF `VAM` scores. 

```{r}
caf <- subset(fibro, subset = label %in% c("iCAF", "myCAF", "apCAF"))
p26a <- data.frame(t(caf@assays$VAMcdf@data)) %>% 
        bind_cols((caf@meta.data %>% dplyr::select(label))) %>% 
        ggplot(aes(x = label, y = apCAF, fill = label)) + 
        geom_violin(scale = "width", draw_quantiles = 0.5, size = 1.2) + 
        geom_signif(comparisons = list(c("apCAF", "iCAF"), c("apCAF", "myCAF")), size = 1.2, textsize = 8, margin_top = .075, 
                    test = wilcox.test, color = "black", map_signif_level = TRUE, step_increase = .15) +
        scale_fill_manual(values = paletteer_d("ggsci::default_nejm")[c(5, 4, 2)]) +
        scale_y_continuous(breaks = seq(0, 1, 0.25), limits = c(0, 1.29)) + 
        labs(y = "Elyada apCAF", x = NULL, fill = NULL) + 
        theme_minimal() + 
        theme(panel.grid = element_blank(), 
              panel.border = element_rect(fill = NA, size = 1), 
              legend.position = "none", 
              axis.title.y = element_text(size = 24), 
              axis.text = element_text(size = 22), 
              axis.text.x = element_text(size = 22, angle = 45, vjust = 1, hjust = 1), 
              axis.ticks = element_line(), 
              plot.subtitle = element_text(face = "italic", size = 10))
p26b <- data.frame(t(caf@assays$VAMcdf@data)) %>% 
        bind_cols((caf@meta.data %>% dplyr::select(label))) %>% 
        ggplot(aes(x = label, y = iCAF, fill = label)) + 
        geom_violin(scale = "width", draw_quantiles = 0.5, size = 1.2) + 
        geom_signif(comparisons = list(c("iCAF", "apCAF"), c("iCAF", "myCAF")), size = 1.2, textsize = 8, margin_top = .075, 
                    test = wilcox.test, color = "black", map_signif_level = TRUE, step_increase = .15) +
        scale_fill_manual(values = paletteer_d("ggsci::default_nejm")[c(5, 4, 2)]) +
        scale_y_continuous(breaks = seq(0, 1, 0.25), limits = c(0, 1.29)) + 
        labs(y = "Elyada iCAF", x = NULL, fill = NULL) + 
        theme_minimal() + 
        theme(panel.grid = element_blank(), 
              panel.border = element_rect(fill = NA, size = 1), 
              legend.position = "none", 
              axis.title.y = element_text(size = 24), 
              axis.text = element_text(size = 22), 
              axis.text.x = element_text(size = 22, angle = 45, vjust = 1, hjust = 1), 
              axis.ticks = element_line(), 
              plot.subtitle = element_text(face = "italic", size = 10))
p26c <- data.frame(t(caf@assays$VAMcdf@data)) %>% 
        bind_cols((caf@meta.data %>% dplyr::select(label))) %>% 
        ggplot(aes(x = label, y = myCAF, fill = label)) + 
        geom_violin(scale = "width", draw_quantiles = 0.5, size = 1.2) + 
        geom_signif(comparisons = list(c("myCAF", "iCAF"), c("myCAF", "apCAF")), size = 1.2, textsize = 8, margin_top = .075, 
                    test = wilcox.test, color = "black", map_signif_level = TRUE, step_increase = .15) +
        scale_fill_manual(values = paletteer_d("ggsci::default_nejm")[c(5, 4, 2)]) +
        scale_y_continuous(breaks = seq(0, 1, 0.25), limits = c(0, 1.29)) + 
        labs(y = "Elyada myCAF", x = NULL, fill = NULL) + 
        theme_minimal() + 
        theme(panel.grid = element_blank(), 
              panel.border = element_rect(fill = NA, size = 1), 
              legend.position = "none", 
              axis.title.y = element_text(size = 24), 
              axis.text = element_text(size = 22), 
              axis.text.x = element_text(size = 22, angle = 45, vjust = 1, hjust = 1), 
              axis.ticks = element_line(), 
              plot.subtitle = element_text(face = "italic", size = 10))
p26d <- data.frame(t(caf@assays$VAMcdf@data)) %>% 
        bind_cols((caf@meta.data %>% dplyr::select(label))) %>% 
        ggplot(aes(x = label, y = `Pan.CAF`, fill = label)) + 
        geom_violin(scale = "width", draw_quantiles = 0.5, size = 1.2) +
        scale_fill_manual(values = paletteer_d("ggsci::default_nejm")[c(5, 4, 2)]) +
        labs(y = "Elyada Pan-CAF", x = NULL, fill = NULL) + 
        theme_minimal() + 
        theme(panel.grid = element_blank(), 
              panel.border = element_rect(fill = NA, size = 1), 
              legend.position = "none", 
              axis.title.y = element_text(size = 24), 
              axis.text = element_text(size = 22), 
              axis.text.x = element_text(size = 22, angle = 45, vjust = 1, hjust = 1), 
              axis.ticks = element_line(), 
              plot.subtitle = element_text(face = "italic", size = 10))
```

The `VAM` score distributions differ significantly between our labelled clusters. 

```{r, fig.width=3, fig.height=4.8}
p26e <- (p26a | p26b)/ (p26c | p26d)
p26e
```

## Ductal Cells

### Recluster Ductal Cells

We use KRT8 expression to show the ductal cells residing in clusters 5 and 9. We note that some regions of clusters 5 and 9 have no KRT8 expression, which likely means that they are composed of different cell types. 

```{r}
p27 <- FeaturePlot(pdac, reduction = "tsne", features = "KRT8", pt.size = 0.75) + 
       scale_color_gradientn(colors = wesanderson::wes_palette("Zissou1")) + 
       theme_yehlab() + 
       NoLegend() + 
       theme(axis.title = element_blank())
p27 / p4
```

These are the cells we'll be reclustering. 

```{r}
ductal_cells <- rownames(pdac@meta.data[pdac@meta.data$seurat_clusters %in% c(5, 9), ])
p28 <- DimPlot(pdac, reduction = "tsne", cells.highlight = ductal_cells, cols.highlight = "navy", pt.size = 0.75) + 
       labs(x = "Fit-SNE 1", y = "Fit-SNE 2") + 
       theme_yehlab() + 
       NoLegend()
p28 / p3
```

We run `SCISSORS` to recluster the cells. 

```{r}
ductal <- ReclusterCells(pdac, 
                         which.clust = c(5, 9), 
                         merge.clusters = TRUE, 
                         n.HVG = 4000, 
                         n.PC = 20, 
                         k.vals = c(20, 30, 40, 50), 
                         resolution.vals = c(.2, .3, .4), 
                         nn.metric = "euclidean", 
                         redo.embedding = TRUE, 
                         random.seed = 629)
DimPlot(ductal)
duct_pc <- Embeddings(ductal, reduction = "pca")
```

Again, we'll run Fit-SNE on the reclustered cells. 

```{python}
# import data
duct_pc = r.duct_pc
# run Fit-SNE
affin_duct = PerplexityBasedNN(duct_pc, perplexity=40, random_state=629)
init = initialization.pca(duct_pc, random_state=629)
tsne_duct = TSNEEmbedding(init, affin_duct, negative_gradient_method='fft')
embed_d1 = tsne_duct.optimize(n_iter=350, exaggeration=15, momentum=0.9)
embed_d2 = embed_d1.optimize(n_iter=750, exaggeration=1, momentum=0.4)
```

We pull the results into R, making sure to save the Barnes-Hut t-SNE results in another reduction slot.

```{r}
embed_duct <- as.matrix(py$embed_d2)
rownames(embed_duct) <- colnames(ductal)
ductal@reductions$bh_tsne <- ductal@reductions$tsne
ductal@reductions$tsne<- CreateDimReducObject(embeddings = embed_duct, 
                                              key = "FitSNE_", 
                                              assay = "SCT", 
                                              global = TRUE)
p29 <- DimPlot(ductal, reduction = "tsne", pt.size = 1.5) + 
       scale_color_manual(values = paletteer_d("ggsci::default_nejm")) + 
       labs(x = "Fit-SNE 1", y = "Fit-SNE 2") + 
       theme_yehlab() + 
       guides(color = guide_legend(nrow = 1, override.aes = list(size = 4)))
p29
```

### Cell Type Identification

#### Lipid Processing

We use ANPEP & FABP1 expression to identify the lipid processing ductal cells in cluster 5. 

```{r}
p30 <- FeaturePlot(ductal, reduction = "tsne", features = "ANPEP", pt.size = 1.5) + 
       scale_color_gradientn(colors = wesanderson::wes_palette("Zissou1")) + 
       theme_yehlab() + 
       NoLegend() + 
       theme(axis.title = element_blank())
p31 <- FeaturePlot(ductal, reduction = "tsne", features = "FABP1", pt.size = 1.5) + 
       scale_color_gradientn(colors = wesanderson::wes_palette("Zissou1")) + 
       theme_yehlab() + 
       NoLegend() + 
       theme(axis.title = element_blank())
(p30 | p31) / p29
```

#### Secretory

Expression of SOD3 and CFTR reveals the secretory cells in cluster 2.

```{r}
p32 <- FeaturePlot(ductal, reduction = "tsne", features = "SOD3", pt.size = 1.5) + 
       scale_color_gradientn(colors = wesanderson::wes_palette("Zissou1")) + 
       theme_yehlab() + 
       NoLegend() + 
       theme(axis.title = element_blank())
p33 <- FeaturePlot(ductal, reduction = "tsne", features = "CFTR", pt.size = 1.5) + 
       scale_color_gradientn(colors = wesanderson::wes_palette("Zissou1")) + 
       theme_yehlab() + 
       NoLegend() + 
       theme(axis.title = element_blank())
(p32 | p33) / p29
```

#### Acinar 

We show that cluster 3 is composed of acinar cells using its high CTRB2 expression.

```{r}
p34 <- FeaturePlot(ductal, reduction = "tsne", features = "CTRB2", pt.size = 1.5) + 
       scale_color_gradientn(colors = wesanderson::wes_palette("Zissou1")) + 
       theme_yehlab() + 
       NoLegend() + 
       theme(axis.title = element_blank())
p34 / p29
```

#### Basal-like & Classical PDAC

We use TFF1 and TFF2 expression to annotate the classical 1 epithelial cells in clusters 0 and 1. 

```{r}
p35 <- FeaturePlot(ductal, reduction = "tsne", features = "TFF1", pt.size = 1.5) + 
       scale_color_gradientn(colors = wesanderson::wes_palette("Zissou1")) + 
       theme_yehlab() + 
       NoLegend() + 
       theme(axis.title = element_blank())
p36 <- FeaturePlot(ductal, reduction = "tsne", features = "TFF2", pt.size = 1.5) + 
       scale_color_gradientn(colors = wesanderson::wes_palette("Zissou1")) + 
       theme_yehlab() + 
       NoLegend() + 
       theme(axis.title = element_blank())
(p35 | p36) / p29
```

The Classical 2 marker CRISP3 is exclusively expressed in cluster 6. We also see that the canonical classical PDAC marker GATA6 is lightly expressed across the Classical PDAC clusters (0, 1, 6). 

```{r}
p37 <- FeaturePlot(ductal, reduction = "tsne", features = "CRISP3", pt.size = 1.5) + 
       scale_color_gradientn(colors = wesanderson::wes_palette("Zissou1")) + 
       theme_yehlab() + 
       NoLegend() + 
       theme(axis.title = element_blank())
p38 <- FeaturePlot(ductal, reduction = "tsne", features = "GATA6", pt.size = 1.5) + 
       scale_color_gradientn(colors = wesanderson::wes_palette("Zissou1")) + 
       theme_yehlab() + 
       NoLegend() + 
       theme(axis.title = element_blank())
(p37 | p38) / p29
```

The basal compartment weights from `DECODER` are highest in cluster 4, which we will denote as being composed of basal-like PDAC. 

```{r}
p39 <- FeaturePlot(ductal, reduction = "tsne", features = "basal_like", pt.size = 1.5) + 
       scale_color_gradientn(colors = wesanderson::wes_palette("Zissou1")) + 
       labs(title = "Basal") + 
       theme_yehlab() + 
       NoLegend() + 
       theme(axis.title = element_blank())
p40 <- FeaturePlot(ductal, reduction = "tsne", features = "classical", pt.size = 1.5) + 
       scale_color_gradientn(colors = wesanderson::wes_palette("Zissou1")) + 
       labs(title = "Classical") + 
       theme_yehlab() + 
       NoLegend() + 
       theme(axis.title = element_blank())
p41 <- FeaturePlot(ductal, reduction = "tsne", features = "bc_ratio", pt.size = 1.5) + 
       scale_color_gradientn(colors = wesanderson::wes_palette("Zissou1")) + 
       labs(title = "Basal:Classical") + 
       theme_yehlab() + 
       NoLegend() + 
       theme(axis.title = element_blank())
p42 <- FeaturePlot(ductal, reduction = "tsne", features = "malig", pt.size = 1.5) + 
       scale_color_manual(values = wes_palette("Zissou1", n = 5)[c(5, 2)])  + 
       labs(title = "Malignant") + 
       theme_yehlab() + 
       NoLegend() + 
       theme(axis.title = element_blank())
(p39 | p40 | p41 | p42) / p29
```

We'll confirm the Basal-like and Classical PDAC cluster identities by using `VAM` to generate scores for the basal and classical gene sets from Moffitt *et al* (2015):

```{r}
pdac_gene_sets <- list(classical_genes, basal_genes)
names(pdac_gene_sets) <- c("Classical PDAC", "Basal-like PDAC")
for (i in seq(pdac_gene_sets)) {
  pdac_gene_sets[[i]] <- pdac_gene_sets[[i]][pdac_gene_sets[[i]] %in% rownames(ductal)]
}
ductal <- vamForSeurat(ductal, 
                       gene.set.collection = pdac_gene_sets, 
                       gamma = TRUE)
DefaultAssay(ductal) <- "VAMcdf"
```

The Classical PDAC scores are highest in clusters 0 & 1 and very low in cluster 4, while the Basal-like scores are highest in cluster 4. 

```{r}
p43 <- FeaturePlot(ductal, reduction = "tsne", features = "Classical PDAC", pt.size = 1.5) + 
       scale_color_gradientn(colors = wesanderson::wes_palette("Zissou1")) + 
       labs(title = "Classical PDAC") + 
       theme_yehlab() + 
       NoLegend() + 
       theme(axis.title = element_blank())
p44 <- FeaturePlot(ductal, reduction = "tsne", features = "Basal-like PDAC", pt.size = 1.5) + 
       scale_color_gradientn(colors = wesanderson::wes_palette("Zissou1")) + 
       labs(title = "Basal-like PDAC") + 
       theme_yehlab() + 
       NoLegend() + 
       theme(axis.title = element_blank())
(p43 | p44) / p29
```

We assign celltype labels to our ductal cells. 

```{r}
DefaultAssay(ductal) <- "SCT"
ductal@meta.data %<>% mutate(label = case_when(seurat_clusters == 0 ~ "Classical 1", 
                                               seurat_clusters == 1 ~ "Classical 1", 
                                               seurat_clusters == 2 ~ "Secretory", 
                                               seurat_clusters == 3 ~ "Acinar", 
                                               seurat_clusters == 4 ~ "Basal-like", 
                                               seurat_clusters == 5 ~ "Lipid Proc.", 
                                               seurat_clusters == 6 ~ "Classical 2")) 
Idents(ductal) <- "label"
p45 <- DimPlot(ductal, reduction = "tsne",pt.size = 1.5) + 
       scale_color_manual(values = paletteer_d("ggsci::default_nejm")) + 
       labs(x = "Fit-SNE 1", y = "Fit-SNE 2") + 
       theme_yehlab() + 
       theme(plot.title = element_blank()) + 
       guides(color = guide_legend(nrow = 1, override.aes = list(size = 4)))
p45
```

Lastly, we'll also provide some visualizations of canonical markers for basal vs. classical PDAC. 

```{r}
pdac_ductal <- subset(ductal, subset = label %in% c("Classical 1", "Classical 2", "Basal-like"))
pdac_ductal@meta.data %<>% mutate(label_broad = case_when(label == "Basal-like" ~ label, TRUE ~ "Classical"))
p46a <- t(data.frame(pdac_ductal@assays$SCT@data)) %>% 
        as.data.frame() %>% 
        mutate(label = pdac_ductal$label) %>% 
        relocate(label, GATA6) %>% 
        dplyr::select(label, GATA6) %>% 
        ggplot(aes(x = label, fill = label, y = GATA6)) + 
        geom_violin(scale = "width", draw_quantiles = 0.5, size = 1) + 
        geom_signif(test = wilcox.test, map_signif_level = TRUE, step_increase = .15, size = 1.2, textsize = 8, 
                    comparisons = list(c("Classical 1", "Basal-like"), c("Classical 2", "Basal-like"))) + 
        scale_fill_manual(values = paletteer_d("ggsci::default_nejm")[c(5, 1, 4)]) + 
        scale_y_continuous(limits = c(0, 1.77), breaks = c(0, 0.5, 1)) + 
        labs(y = "GATA6", title = NULL, x = NULL) + 
        theme(panel.grid = element_blank(), 
              legend.position = "none", 
              panel.background = element_blank(), 
              panel.border = element_rect(fill = NA, size = 1), 
              axis.title.x = element_text(size = 24), 
              axis.title.y = element_text(size = 24), 
              axis.text.y = element_text(size = 20), 
              axis.text.x = element_text(size = 20, angle = 45, vjust = 1, hjust = 1))
p46b <- ggplot(pdac_ductal@meta.data, aes(x = label, fill = label, y = basal_like)) + 
        geom_violin(scale = "width", draw_quantiles = 0.5, size = 1) +
        geom_signif(test = wilcox.test, map_signif_level = TRUE, step_increase = .15, size = 1.2, textsize = 8, 
                    comparisons = list(c("Classical 1", "Basal-like"), c("Classical 2", "Basal-like"))) + 
        scale_fill_manual(values = paletteer_d("ggsci::default_nejm")[c(5, 1, 4)]) + 
        scale_y_continuous(limits = c(0, 1.3), breaks = c(0, 0.25, 0.5, 0.75, 1)) + 
        labs(y = "DECODER basal-like", title = NULL, x = NULL) + 
        theme(panel.grid = element_blank(), 
              legend.position = "none", 
              panel.background = element_blank(), 
              panel.border = element_rect(fill = NA, size = 1), 
              axis.title.x = element_text(size = 24), 
              axis.title.y = element_text(size = 24), 
              axis.text.y = element_text(size = 20), 
              axis.text.x = element_text(size = 20, angle = 45, vjust = 1, hjust = 1))
p46c <- ggplot(pdac_ductal@meta.data, aes(x = label, fill = label, y = classical)) + 
        geom_violin(scale = "width", draw_quantiles = 0.5, size = 1) +
        geom_signif(test = wilcox.test, map_signif_level = TRUE, step_increase = .15, size = 1.2, textsize = 8, 
                    comparisons = list(c("Classical 1", "Basal-like"), c("Classical 2", "Basal-like"))) + 
        scale_fill_manual(values = paletteer_d("ggsci::default_nejm")[c(5, 1, 4)]) + 
        scale_y_continuous(limits = c(0, 1.25), breaks = c(0, 0.25, 0.5, 0.75, 1)) + 
        labs(y = "DECODER classical", title = NULL, x = NULL) + 
        theme(panel.grid = element_blank(), 
              legend.position = "none", 
              panel.background = element_blank(), 
              panel.border = element_rect(fill = NA, size = 1), 
              axis.title.x = element_text(size = 24), 
              axis.title.y = element_text(size = 24), 
              axis.text.y = element_text(size = 20), 
              axis.text.x = element_text(size = 20, angle = 45, vjust = 1, hjust = 1))
p46d <- ggplot(pdac_ductal@meta.data, aes(x = label, fill = label, y = bc_ratio)) + 
        geom_violin(scale = "width", draw_quantiles = 0.5, size = 1) +
        geom_signif(test = wilcox.test, map_signif_level = TRUE, step_increase = .15, size = 1.2, textsize = 8, 
                    comparisons = list(c("Classical 1", "Basal-like"), c("Classical 2", "Basal-like"))) + 
        scale_fill_manual(values = paletteer_d("ggsci::default_nejm")[c(5, 1, 4)]) + 
        scale_y_continuous(limits = c(0, 6.35)) + 
        labs(y = "DECODER bcRatio", title = NULL, x = NULL) + 
        theme(panel.grid = element_blank(), 
              legend.position = "none", 
              panel.background = element_blank(), 
              panel.border = element_rect(fill = NA, size = 1), 
              axis.title.x = element_text(size = 24), 
              axis.title.y = element_text(size = 24), 
              axis.text.y = element_text(size = 20), 
              axis.text.x = element_text(size = 20, angle = 45, vjust = 1, hjust = 1))
p46e <- t(data.frame(pdac_ductal@assays$VAMcdf@data)) %>% 
        as.data.frame() %>% 
        mutate(label = pdac_ductal@meta.data$label) %>% 
        ggplot(aes(x = label, fill = label, y = `Classical PDAC`)) + 
        geom_violin(scale = "width", draw_quantiles = 0.5, size = 1) +
        geom_signif(test = wilcox.test, map_signif_level = TRUE, step_increase = .15, size = 1.2, textsize = 8, 
                    comparisons = list(c("Classical 1", "Basal-like"), c("Classical 2", "Basal-like"))) + 
        scale_fill_manual(values = paletteer_d("ggsci::default_nejm")[c(5, 1, 4)]) + 
        scale_y_continuous(limits = c(0, 1.3), breaks = c(0, 0.25, 0.5, 0.75, 1)) + 
        labs(y = "Moffitt classical", title = NULL, x = NULL) + 
        theme(panel.grid = element_blank(), 
              legend.position = "none", 
              panel.background = element_blank(), 
              panel.border = element_rect(fill = NA, size = 1), 
              axis.title.x = element_text(size = 24), 
              axis.title.y = element_text(size = 24), 
              axis.text.y = element_text(size = 20), 
              axis.text.x = element_text(size = 20, angle = 45, vjust = 1, hjust = 1))
p46f <- t(data.frame(pdac_ductal@assays$VAMcdf@data)) %>% 
        as.data.frame() %>% 
        mutate(label = pdac_ductal@meta.data$label) %>% 
        ggplot(aes(x = label, fill = label, y = `Basal-like PDAC`)) + 
        geom_violin(scale = "width", draw_quantiles = 0.5, size = 1) +
        geom_signif(test = wilcox.test, map_signif_level = TRUE, step_increase = .15, size = 1.2, textsize = 8, 
                    comparisons = list(c("Classical 1", "Basal-like"), c("Classical 2", "Basal-like"))) + 
        scale_fill_manual(values = paletteer_d("ggsci::default_nejm")[c(5, 1, 4)]) + 
        scale_y_continuous(limits = c(0, 1.3), breaks = c(0, 0.25, 0.5, 0.75, 1)) + 
        labs(y = "Moffitt basal-like", title = NULL, x = NULL) + 
        theme(panel.grid = element_blank(), 
              legend.position = "none", 
              panel.background = element_blank(), 
              panel.border = element_rect(fill = NA, size = 1), 
              axis.title.x = element_text(size = 24), 
              axis.title.y = element_text(size = 24), 
              axis.text.y = element_text(size = 20), 
              axis.text.x = element_text(size = 20, angle = 45, vjust = 1, hjust = 1))
```

The markers / gene sets match up with our cluster labels. 

```{r, fig.width=6.5, fig.height=3}
p47 <- (p46f | p46e | p46b | p46d | p46c | p46c)
p47
```
